---
title: "about.Rmd"
author: "Katherine Wang"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(tidyverse)
library(janitor)
library(ggplot2)
library(tidycensus)

library(ggmap)
library(maps)
library(mapdata)

library(readxl)
library(janitor)
```

```{r cleanData, include = FALSE}

# For each data set, I cleaned its variables' names and added a new column,
# year, in order to facilitate organizing the joined data sets later on. I also
# selected the variables I wanted, under the name "datapoints."

datapoints <- c("geo_geoid", "geo_name", "se_a02001_001", "se_a02001_002", "se_a02001_003", "se_a03001_001", "se_a03001_002", "se_a03001_003", "se_a12003a_001", "se_a12003a_002", "se_a12003a_003", "se_a12003b_001", "se_a12003b_002", "se_a12003b_003", "se_a17005a_001", "se_a17005a_002", "se_a17005a_003", "se_a17005b_001", "se_a17005b_002", "se_a17005b_003", "se_a17006a_001", "se_a17006a_002", "se_a17006a_003", "se_a17006b_001", "se_a17006b_002", "se_a17006b_003")

acs2009 <- read_csv("raw-data/ACS2009.csv") %>% 
  clean_names() %>% 
  mutate(year = "2009") %>% 
  select(datapoints)

acs2010 <- read_csv("raw-data/ACS2010.csv") %>% 
  clean_names() %>% 
  mutate(year = "2010") %>% 
  select(datapoints)

acs2011 <- read_csv("raw-data/ACS2011.csv") %>% 
  clean_names() %>% 
  mutate(year = "2011") %>% 
  select(datapoints)

acs2012 <- read_csv("raw-data/ACS2012.csv") %>% 
  clean_names() %>% 
  mutate(year = "2012") %>% 
  select(datapoints)

acs2013 <- read_csv("raw-data/ACS2013.csv") %>% 
  clean_names() %>% 
  mutate(year = "2013") %>% 
  select(datapoints)

acs2014 <- read_csv("raw-data/ACS2014.csv") %>% 
  clean_names() %>% 
  mutate(year = "2014") %>% 
  select(datapoints)

acs2015 <- read_csv("raw-data/ACS2015.csv") %>% 
  clean_names() %>% 
  mutate(year = "2015") %>% 
  select(datapoints)

acs2016 <- read_csv("raw-data/ACS2016.csv") %>% 
  clean_names() %>% 
  mutate(year = "2016") %>% 
  select(datapoints)

acs2017 <- read_csv("raw-data/ACS2017.csv") %>% 
  clean_names() %>% 
  mutate(year = "2017") %>% 
  select(datapoints)

acs2018 <- read_csv("raw-data/ACS2018.csv") %>% 
  clean_names() %>% 
  mutate(year = "2018") %>% 
  select(datapoints)

# Since this is not part of the ACS data sets, I did not have to add a year
# column to it.

nces <- read_csv("raw-data/NCES.csv") %>% 
  clean_names()

# In order to join all the ACS data sets to each other, I used inner_join. Since
# I only know how to join two data sets to each other at a time, I did this
# repeatedly until all the data sets were joined together. Then, for the final
# joined data set, I got rid of the columns that only had NAs in them.

joined1 <- rbind(acs2009, acs2010)
joined2 <- inner_join(joined1, acs2011, by = "geo_geoid")
joined3 <- inner_join(joined2, acs2012, by = "geo_geoid")
joined4 <- inner_join(joined3, acs2013, by = "geo_geoid")
joined5 <- inner_join(joined4, acs2014, by = "geo_geoid")
joined6 <- inner_join(joined5, acs2015, by = "geo_geoid")
joined7 <- inner_join(joined6, acs2016, by = "geo_geoid")
joined8 <- inner_join(joined7, acs2017, by = "geo_geoid")
joined_final <- inner_join(joined8, acs2018, by = "geo_geoid")

joined_final <- joined_final[, colMeans(is.na(joined_final)) < 1]
```


```{r acsGraph, include = T, echo = F}

# This is for the year 2009 only. I first created two separate tibbles, one
# displaying the male school dropout rate and the other displaying the female
# school dropout rate. To do this, I selected the three variables I needed,
# which were the geoid of the school districts, the total number of male/female
# students enrolled, and the number of male/female students who dropped out. For
# each tibble, I created a new variable, perc_dropout, which measures the
# proportion of students who dropped out compared to the total student
# population. I also created a variable, type, to facilitate with faceting the
# two tables later on. I only selected the first 10 data points for now because
# the program would otherwise take too long to run.

male <- joined_final %>% 
  select(geo_geoid, 
         se_a12003a_001, 
         se_a12003a_002) %>% 
  mutate(perc_dropout = (se_a12003a_002/se_a12003a_001)*100) %>% 
  mutate(type = "male") %>% 
  mutate(total = se_a12003a_001, drop = se_a12003a_002) %>% 
  select(geo_geoid, total, drop, perc_dropout, type) %>% 
  head(10)

female <- joined_final %>% 
  select(geo_geoid, 
         se_a12003b_001, 
         se_a12003b_002) %>% 
  mutate(perc_dropout = (se_a12003b_002/se_a12003b_001)*100) %>%
  mutate(type = "female") %>% 
  mutate(total = se_a12003b_001, drop = se_a12003b_002) %>% 
  select(geo_geoid, total, drop, perc_dropout, type) %>% 
  head(10)

# I combined the two tibbles with rbind. To do this, I had to ensure that all of
# their variable names matched, and that I could distinguish between the data
# sets with the "type" variable, which shows whether or not the data is about
# male or female students.

drop <- rbind(male, female)
  
# I turned this into a bar chart, with the school districts on the x axis and
# the percent dropout rate on the y axis. I gave the graph its respective tites
# and axes labels. Finally, I used facet wrap to place the female and male
# dropout rates side by side.

ggplot(drop, aes(x = geo_geoid, y = perc_dropout)) +
  geom_col() +
  labs(title = "Difference in School Dropout Rates Based on Sex", 
       subtitle = "For Population Ages 16-19",
       caption = "Source: American Communities Survey",
       x = "School District ID",
       y = "Dropout Rate") +
  facet_wrap(~ type) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r getData, include = F}
# I obtained data which depicts the differences in earnings for men and women
# from the Status of Women in the States. I cleaned this data and changed its
# state names to lowercase to match those in the "states" data, which will
# appear in the next r chunk. This is so that the two datasets can be combined
# more easily.

earnings <- read_xlsx("raw-data/EarningRatios.xlsx") %>% 
  clean_names() %>% 
  mutate(region = state)

earnings$region <- tolower(earnings$state)
```


```{r map, include = T, echo = F}
# I downloaded data for the states, which is already stored in R. 

states <- map_data("state")

# I created a ggplot of the states just to see how the map would look like
# before I added my own data to it.

p <- ggplot(data = states, aes(x = long, 
                   y = lat, group = group)) + 
  geom_polygon(fill = "lightgray", 
               color = "white") + 
  coord_fixed(1.3) +
  guides(fill = FALSE)

# I then combined the two data sets, states and earnings, so that I can graph
# earnings by state.

combined <- left_join(states, earnings, by = "region")

# I created a map showing earning ratios by state. The x and y axis are the map
# coordinates, and I filled the graph with a variable called percent in the
# earnings data set, which depicts the percent of women's earnings in comparison
# to men's.

ggplot(combined, aes(x = long, y = lat, group = group, fill = percent)) +
  geom_polygon(color = "white") +
  coord_fixed(1.3) +
  labs(title = "Gender Earnings Ratio for Men and Women with a Bachelor's Degree or Higher", 
       subtitle = "In 2013",
       caption = "Source: Status of Women in the States",
       fill = "Earnings Ratio") +
  theme_void()
```


**Background**


The goal of this project is to discover the extent to which a school’s inequality mirrors that of its community’s. Although some believe that schools act as the springboard of social equality and progress, others argue that schools merely reinforce the existing stratification found in their societies. In order to test the idea of whether schools serve as social mirrors, I will analyze two main forms of inequality, race and gender, in two separate spheres -- the school and the community. Ultimately, I hope to discover whether a correlative relationship exists between inequality in a community and inequality in schools.


I gathered my data from two sources: the American Community Surveys (ACS) from the US Census Bureau and the Status of Women in the States. From ACS’ data, I extracted information about each school district’s distribution of race and gender, unemployment rate by race and gender, average household income by race, and school dropout rate by sex. From the data from the Status of Women in the States, I was able to analyze the gender gap in earnings to try to evaluate a community's level of gender inequality. Although I downloaded data from the National Center for Education Statistics (NCES), which included data on school enrollments by race and sex, I am not sure if I will be implementing it as of now.


Currently, I have created a graphic displaying levels of gender inequality in communities, as well as a chart showing gender inequality in schools in terms of dropout rates. I will create more graphics regarding racial inequality, and I hope to find a way to combine the different graphs I make. 


Url to my repo: https://github.com/katherinewang122/Final-Project-Milestone-3.git