---
title: "about.Rmd"
author: "Katherine Wang"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(tidyverse)
library(janitor)
```

```{r cleanData, include = FALSE}

# For each data set, I cleaned its variables' names and added a new column,
# year, in order to facilitate organizing the joined data sets later on.

acs2009 <- read_csv("raw-data/ACS2009.csv") %>% 
  clean_names() %>% 
  mutate(year = "2009")

acs2010 <- read_csv("raw-data/ACS2010.csv") %>% 
  clean_names() %>% 
  mutate(year = "2010")

acs2011 <- read_csv("raw-data/ACS2011.csv") %>% 
  clean_names() %>% 
  mutate(year = "2011")

acs2012 <- read_csv("raw-data/ACS2012.csv") %>% 
  clean_names() %>% 
  mutate(year = "2012")

acs2013 <- read_csv("raw-data/ACS2013.csv") %>% 
  clean_names() %>% 
  mutate(year = "2013")

acs2014 <- read_csv("raw-data/ACS2014.csv") %>% 
  clean_names() %>% 
  mutate(year = "2014")

acs2015 <- read_csv("raw-data/ACS2015.csv") %>% 
  clean_names() %>% 
  mutate(year = "2015")

acs2016 <- read_csv("raw-data/ACS2016.csv") %>% 
  clean_names() %>% 
  mutate(year = "2016")

acs2017 <- read_csv("raw-data/ACS2017.csv") %>% 
  clean_names() %>% 
  mutate(year = "2017")

acs2018 <- read_csv("raw-data/ACS2018.csv") %>% 
  clean_names() %>% 
  mutate(year = "2018")

# Since this is not part of the ACS data sets, I did not have to add a year
# column to it.

nces <- read_csv("raw-data/NCES.csv") %>% 
  clean_names()

# In order to join all the ACS data sets to each other, I used str_detect to
# find their common term, "ACS," in the data sets. The for loop to allows me to
# join each data set at a time because merge.data.frame only works with joining
# two data frames to each other. I joined them by their common variable,
# Geo_GEOID.

test <- list.files('raw-data', full.names = T)
test <- test[str_detect(test, 'ACS')]

storage <- read_csv(test[1])
for(i in 2:length(test)){
  storage <- merge.data.frame(storage, read_csv(test[i]), by = "Geo_GEOID")
}

aaa <- read_csv(test[1])
bbb <- read_csv(test[2])

aaa <- aaa[, colMeans(is.na(aaa)) < 1]
```


**Background**


The goal of this project is to discover the extent to which a school’s inequality mirrors that of its community’s. Although some believe that schools act as the springboard of social equality and progress, others argue that schools merely reinforce the existing stratification found in their societies. In order to test the idea of whether schools serve as social mirrors, I will analyze two main forms of inequality, race and gender, in two separate spheres -- the school and the community. Ultimately, I hope to discover whether a correlative relationship exists between inequality in a community and inequality in schools.


I gathered my data from two sources: the American Community Surveys (ACS) from the US Census Bureau and the National Center for Education Statistics (NCES). From ACS’ data, I extracted information about each school district’s distribution of race and gender, unemployment rate by race and gender, average household income by race, and school dropout rate by sex. From NCES, I was able to find the data on school enrollments by race and sex.


Currently, I have gathered most of my data and converted them into csv files. I have not yet finished processing my data because I am still trying to find a way to either organize the NCES data by school districts as opposed to agencies or states or find completely new data that is already grouped by school districts. I am also in the process of combining the ACS files into one comprehensive spreadsheet from 2009-2018 because, right now, each spreadsheet represents a different year.


Url to my repo: https://github.com/katherinewang122/Final-Project-Milestone-3.git