---
title: "about.Rmd"
author: "Katherine Wang"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(tidyverse)
library(janitor)
```

```{r cleanData, include = FALSE}
acs2009 <- read_csv("raw-data/ACS2009.csv")
acs2010 <- read_csv("raw-data/ACS2010.csv")
acs2011 <- read_csv("raw-data/ACS2011.csv")
acs2012 <- read_csv("raw-data/ACS2012.csv")
acs2013 <- read_csv("raw-data/ACS2013.csv")
acs2014 <- read_csv("raw-data/ACS2014.csv")
acs2015 <- read_csv("raw-data/ACS2015.csv")
acs2016 <- read_csv("raw-data/ACS2016.csv")
acs2017 <- read_csv("raw-data/ACS2017.csv")
acs2018 <- read_csv("raw-data/ACS2018.csv")
nces <- read_csv("raw-data/NCES.csv")
```


**Background**


The goal of this project is to discover the extent to which a school’s inequality mirrors that of its community’s. Although some believe that schools act as the springboard of social equality and progress, others argue that schools merely reinforce the existing stratification found in their societies. In order to test the idea of whether schools serve as social mirrors, I will analyze two main forms of inequality, race and gender, in two separate spheres -- the school and the community. Ultimately, I hope to discover whether a correlative relationship exists between inequality in a community and inequality in schools.


I gathered my data from two sources: the American Community Surveys (ACS) from the US Census Bureau and the National Center for Education Statistics (NCES). From ACS’ data, I extracted information about each school district’s distribution of race and gender, unemployment rate by race and gender, average household income by race, and school dropout rate by sex. From NCES, I was able to find the data on school enrollments by race and sex.


Currently, I have gathered most of my data and converted them into csv files. I have not yet processed my data because I am still trying to find a way to organize the NCES data by school districts as opposed to agencies or states. I am also in the process of combining the ACS files into one comprehensive spreadsheet from 2009-2018 because, right now, each spreadsheet represents a different year.


Url to my repo: https://github.com/katherinewang122/Final-Project-Milestone-3.git